Decorrelating Errors in Quantum Gates by Random Gate Synthesis

### 2018-07-19 decorrelation/decorrelation.tex
Adding in experimental results, and more references and comments, still very rough

%Thresholds for fault-tolerant quantum computation are often calculated assuming a noise model in which errors are uncorrelated. While convenient for simulation, these error models are often unphysical.  Recent work by Preskill and others has shown that the arbitrarily long computations may be performed even in the presence of spatial correlation, provided the correlation is sufficiently weak and decays sufficiently quickly with distance, but at the cost of a significantly lower threshold. The success of algebraic decorrelation methods, such as dynamical decoupling, demonstrate that quantum control techniques are capable of reducing temporal noise correlations. We propose to introduce similar methods to effect the spatial decorrelation of errors in quantum circuits, thereby increasing the threshold for fault-tolerant computation in such systems.

@@

Another assumption that is instrumental in proving many threshold theorems is Markovianity \cite{Kitaev1997}. There are many ways in which devices can behave in non-Markovian ways, and NISQ devices have already been shown to be subject to drift in controls, \cite{Kelly2018} and in general non-Markovian noise \cite{BlumeKohout2017}. Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit composition to try to remove correlated noise. [WHY IS WHAT WE'RE DOING BETTER?]

@@

For example, randomized benchmarking and tomography will report incorrect answers without any syndrome, [CITATION NEEDED], and gateset tomography will report that the gateset failed to be Markovian, but will fail to diagnose in what way it was non-Markovian. Because BCSs change the error to a depolorizing channel, the correlations in the noise can be massively reduced, and the process can be made much closer to Markovian, at no real cost to the fidelity of the implemented gate.

@@

\section{1Q Gates}\label{1Q Gates}
 In this section, we present numerical results on generating one-qubit gates that together with $RZ(\theta$) rotations are universal for one-qubit computation. Our control Hamiltonian is given as: 
\begin{equation}\label{eq:1Qham}
  H = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
$\epsilon$ is a normally distributed random variable, centered at $0$ with standard deviation $.001$, corresponding to a time varying Stark shift on the qubit. $\delta$ is also a normally distributed random variable with the same parameters, and is meant to represent random fluctuations in the control electronics, such as diurnal variations resulting from temperature instability. We assume that the errors on $\sigma_x$ and $\sigma_y$ are perfecly correlated, as mentioned in Section \ref{ocp}.
[ADD ANALYSIS OF PLOTS]
\begin{figure*}
\centering
\includegraphics[width=\textwidth]{x2.png}
\caption{In the above we have used our method to produce a BCS for a $RX(\frac{\pi}{2})$ pulse, with the Hamiltonian given in \ref{eq:1Qham}. A) shows the sum of the squared norms of the off diagonal terms in the Pauli Transfer Matrix of $\mathcal{E}$. B) gives the corresponding infidelity for each member of the BCS, and the optimal choice of weights.}
\end{figure*}

@@

\section{2Q Gates}\label{2Q Gates}
 In this section, we present numerical results on a generating one and two-qubit gates that together with $RZ(\theta)$ rotations are universal for two-qubit computation. Our control Hamiltonian is given as: 
\begin{equation} \label{eq:2Qham}
H = \exp{(-i\frac{\sigma_z^1\otimes\sigma_z^2}{4})} + \sum_{j=1}^2(\epsilon_j\sigma_z^j + (1 + \delta_j)(c_x^jx(t)\sigma_x^j + c_y^j(t)\sigma_y^j))
\end{equation}
We again assume that the standard deviations are $.001$ on all parameters $\delta_j$ and $c^j_i$ and that errors on the single qubit $\sigma_x$ and $\sigma_y$ rotations are perfectly correlated.
[ADD ANALYSIS OF PLOTS]
%These plots take a little while to make, so I can't iterate *that* fast - clearly the fonts need to be bigger, and there should be a title what else add a legend saying the black is the optimal?
\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{2Q.png}
\caption{In the above we have used our method to produce a BCS for a ZZ($\frac{\pi}{2}$) entangling gate with controls given by \ref{eq:2Qham}. A) shows the sum of the squared norms of the off diagonal terms in the Pauli Transfer Matrix of $\mathcal{E}$. B) gives the corresponding infidelity for each member of the BCS, and the optimal choice of weights.}
\end{figure*}

>>>>

@@

Another assumption that is instrumental in proving many threshold theorems is Markovianity \cite{Kitaev1997}. There are many ways in which devices can behave in non-Markovian ways, and NISQ devices have already been shown to be subject to drift in controls, \cite{Kelly2018} and in general non-Markovian noise \cite{BlumeKohout2017}. Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit composition to try to remove correlated noise. Indeed, other have even implemented these routines natively in hardware \cite{Ware2018}. These routines all require some level of sophistication. In the case of \cite{Campbell2017}, the analysis assumes Solovay-Kitaev [CITATION]-like decomposition of unitaries, and treats Pauli and Clifford gates as being free to implement, which is far from the reality of modern quantum computers. In \cite{Wallman2016} and the implementation in \cite{Ware2018}, the random frames are implemented in each sequence by precompiling the inserted Pauli frame rotations with the Clifford operations. In general, this cannot be done efficiently for non-Clifford operations, and in addition, this introduces a linear-time precompilation overhead for shot of the circuit.

@@

For example, randomized benchmarking and tomography will report incorrect answers without any syndrome, [CITATION NEEDED], and gateset tomography will report that the gateset failed to be Markovian, but will fail to diagnose in what way it was non-Markovian. Because BCSs change the error to a depolorizing channel, the correlations in the noise can be massively reduced, and the process can be made much closer to Markovian, at no real cost to the fidelity of the implemented gate. Unlike \cite{Ware2018}, \cite{Wallman2016}, and \citep{Campbell2017}, this routine can be implemented in the firmware of the quantum computer, and requires only an instruction for a random choice - no pre or post processing is required.

@@

\section{Numerical Results}\label{numerical}
\subsection{1Q Gates}\label{1Q Gates}
 In this section, we present numerical results on generating one-qubit gates that together with $RZ(\theta$) rotations are universal for one-qubit computation. Our control Hamiltonian is given as: 
\begin{equation}\label{eq:1Qham}
  H = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
$\epsilon$ is a normally distributed random variable, centered at $0$ with standard deviation $.001$, corresponding to a time varying Stark shift on the qubit. $\delta$ is also a normally distributed random variable with the same parameters, and is meant to represent random fluctuations in the control electronics, such as diurnal variations resulting from temperature instability. We assume that the errors on $\sigma_x$ and $\sigma_y$ are perfecly correlated, as mentioned in Section \ref{ocp}.
[ADD ANALYSIS OF PLOTS]
\begin{figure*}
\centering
\includegraphics[width=\textwidth]{x2.png}
\caption{In the above we have used our method to produce a BCS for a $RX(\frac{\pi}{2})$ pulse, with the Hamiltonian given in \ref{eq:1Qham}. A) shows the sum of the squared norms of the off diagonal terms in the Pauli Transfer Matrix of $\mathcal{E}$. B) gives the corresponding infidelity for each member of the BCS, and the optimal choice of weights.}
\end{figure*}

@@

\subsection{2Q Gates}\label{2Q Gates}
 In this section, we present numerical results on a generating one and two-qubit gates that together with $RZ(\theta)$ rotations are universal for two-qubit computation. Our control Hamiltonian is given as: 
\begin{equation} \label{eq:2Qham}
H = \exp{(-i\frac{\sigma_z^1\otimes\sigma_z^2}{4})} + \sum_{j=1}^2(\epsilon_j\sigma_z^j + (1 + \delta_j)(c_x^jx(t)\sigma_x^j + c_y^j(t)\sigma_y^j))
\end{equation}
We again assume that the standard deviations are $.001$ on all parameters $\delta_j$ and $c^j_i$ and that errors on the single qubit $\sigma_x$ and $\sigma_y$ rotations are perfectly correlated.
[ADD ANALYSIS OF PLOTS]
%These plots take a little while to make, so I can't iterate *that* fast - clearly the fonts need to be bigger, and there should be a title what else add a legend saying the black is the optimal?
\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{2Q.png}
\caption{In the above we have used our method to produce a BCS for a ZZ($\frac{\pi}{2}$) entangling gate with controls given by \ref{eq:2Qham}. A) shows the sum of the squared norms of the off diagonal terms in the Pauli Transfer Matrix of $\mathcal{E}$. B) gives the corresponding infidelity for each member of the BCS, and the optimal choice of weights.}
\end{figure*}

\section{Experimental Results}\label{experimental}
To demonstrate the realizability of this routine, we implemented it on a superconducting qubit, to calibrate an $RX(\frac{\pi}{2})$ pulse. The qubit being used had a 20.5 $\mu$s T1, and a 11.00 $\mu$s T2, and the pulse shapes being considered were 30ns long top hat pulse sequences, generated via a set-and-hold signal generator using the amplitudes shown in \ref{fig:pulseshapes}. 
\begin{figure}[H]
  \centering
  \includegraphics[width=.75\columnwidth]{pulseshapes.png}
  \caption{The pulseshapes used in the routine. The original pulse was generated by iteratively applying $RX(\frac{\pi}{2})$ pulses, fitting the resulting sinusoid, and updating the controls. The variants were generated via optimal control, seeding the GRAPE algorithm with the initial pulse.}
  \label{fig:pulseshapes}
\end{figure}

In this example we generated 10 candidate pulses using the GRAPE algorithm with a high convergence tolerance, with the intention that the optimizer would find solutions with some degree of noticeable coherent error. For each pulseshape, which enacts some unitary $U_i$, we performed an RPE-like experiment [REF RPE] where the $\sigma_y$ and $\sigma_z$ expectations values were measured at the end of a sequence of $n$ applications of $U_i$, where n ranged from 1 to 100, and each circuit was sampled 1000 times. The results are plotted in [REF SUBPLOTS OF \ref{fig:rebit}]. Then, using the same optimization tool chain as in \ref{numerical}, we picked out the coeffients $p_i$ such that $p_i, U_i$ are a BCS, and performed the same experiment, but where for each shot of the experiment, we precompiled a different circuit. In particular, in our sampling procedure, circuit $U_{i_0}, U_{i_1}, ..., U_{i_{n-1}}$ occurs with probability $p_{i_0}p_{i_1}...p_{i_{n-1}}$. While in this example high fidelity, low coherent error controls can be produced [REFERENCE SUBPLOT OF \ref{fig:pulseshapes}], this demonstrates that this method can reduce coherent errors drastically.

\begin{figure}[H]
  \centering
  \includegraphics[width=.75\columnwidth]{example.png}
  \caption{An experimental implementation of our routine. We should include an array of images, maybe similar to this one, and then comment that in the image representing the BCS (above) that the noise is dominantly incoherent. The colors, ranging from purple to red, correspond to the depth into the sequence, where the darkest purple point is at the top of the rebit at $\ket{0}$. As the color becomes more red, the gate is applied more times, and consequently in the BCS example, moves the qubit's state vector closer to the maximally mixed state at the origin. The angle was computed by taking the arctangent of the $\sigma_y$ and $\sigma_z$ expectations, while the ther radius was computed by adding the two expectations in quadrature.}
  \label{fig:rebit}
\end{figure}

\section{Conclusion}
We have shown numerically and experimentally that drawing from a collection of implementations of a gate with the correct probabilties can reduce coherent error by an order of magnitude, while increasing incoherent error by a smaller amount, at virtually no cost to gate fidelity. We have demonstrated that these approximate controls can be generated through optimal control, and that the minimization problem is tractable. In addition, we have shown that it is possible to perform the routine on existing quantum hardware. Future directions for this work include moving the random gate selection from a precompilation step onto runtime FPGA logic, investigating other optimzation routines such as [GOAT, CRAB], and using more precise benchmarking routines such as GST [REF GST] to more quantitatively investigate the performance of these routines. The code used in this paper is available at [REF MY GITHUB].
